---
title: "Info550_Zelin Wang"
author: "Zelin Wang"
date: "2021/11/9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The cohort consists of patients who initially pick up from the clinic pharmacy from January 2016 to June 2020 for one-year follow up.
We seek to assess the association of the outcome of interest which is viral load undetectability (defined as viral load less than 200 copies/mL) and various variables: pharmacy antiretroviral(ART) pickup frequency at the clinic, age, sex, race, Cd4 and viral load count at initiation of ART, year of ART initial pick up at the clinic. 
Using R markdown, I will clean, sort, code the provided deidentified data set  for analysis 
Association between variables and outcome of interest will be assessed using chi square test  for categorical variables and t-test for continuous variable .
A multivariate model using proc logistic will be used to derive the final model of association between significant variables and the outcome of interest. 

```{r}
renv::init()
```

```{r}
here::i_am("code/Info550_Zelin_Wang.Rmd")
```



#Read in the HIV pharmacy pickup data file and print.
```{r}
pickup<-read.csv(here::here("data/hiv_pickup.csv"),header = TRUE, stringsAsFactors = FALSE)
print(head(pickup,10))
```

```{r}
#Data cleaning
#Sort the raw dataset by sid and DATE_DISPENSED
library(dplyr)
pickup1<-pickup%>%arrange(pickup$sid,as.Date(pickup$DATE_DISPENSED,format="%Y/%m/%d"))
print(head(pickup1,10))
```
```{r}
#Since we are only interested in the one year follow up period, we only need to keep the records within one year since the first pharmacy pick up date.
#In order to fulfill the agenda above, first we need to find out the first pick up date for all sid(Individuals)
library(dplyr)
pickup_initial <- pickup1 %>% group_by(pickup1$sid) %>% filter(row_number()==1)
pickup_initial <- pickup_initial[-c(4)]
print(head(pickup_initial,10))
```
```{r}
#Get the year of the first pickup date for each sid
#Get the frequency of pickup_initial
pickup_initial$year_start=format(as.Date(pickup_initial$DATE_DISPENSED, format="%Y/%m/%d"),"%Y")
print(pickup_initial)
pickup_initial_count<-count(pickup_initial,year_start)
pickup_initial_count<-pickup_initial_count[-7,]
print(pickup_initial_count)
```
```{r}
barp<-barplot(height=pickup_initial_count$n, names=pickup_initial_count$year_start, main = "# of individuals Initial pick up year", col = rainbow(7),ylim = c(0, 700))
text(barp, pickup_initial_count$n+30, labels = pickup_initial_count$n)
```



```{r}
#Add entire one year to the existing date column to obtain one-year periods.
lt <- as.POSIXlt(as.Date(pickup_initial$DATE_DISPENSED,format="%Y/%m/%d"))
library(lubridate)
lt<-lt %m+% months(11)
pickup_initial$date_disp_1YR_FU=as.Date(lt, format="%Y/%m/%d")
print(head(pickup_initial,10))
```

